---
title: "HILL Biological Opinion analysis"
output: html_notebook
---

```{r}
ifelse(Sys.info()[1] == 'Linux',
       source('~/Documents/R/TomosFunctions.R'),
       source('~/R/TomosFunctions.R'))

library(tidyverse)
library(ggplot2)

```

This document contains various analysis used for modeling turtle populations for the Biological Opinion on the Hawaii shallow-set longline fishery. 

#Van Houtan analysis
I'm not convinced about what Kyle did with respect to the "climate-forcing PVA". It appears that he used the standardized log anomaly as the response and explanatory variables.

Looking at leatherback data first.
```{r}

Dc_nests <- read.csv(file = "data/LeatherbackAdjustedNumberNests.csv") %>%
  filter(Year > 1984)
PDO <- read.csv(file = "data/PDO.csv", header = FALSE, na.strings = "NaN")

colnames(PDO) <- c("Year", 1:12)
PDO$avg <- rowMeans(PDO[,2:13])
PDO_Dc <- filter(PDO, (Year >= min(Dc_nests$Year)-22) & 
                   (Year <= max(Dc_nests$Year)-22))

UW36N <- read.csv(file = "data/UW36N_19670101_20170130_Daily_data_int.csv", 
                  header = TRUE, na.strings = "NA")

UW36N %>% filter(month > 5 & month < 9) %>%
  group_by(year) %>%
  summarise(UW_avg = mean(UWint)) %>%
  mutate(log.UW_avg = log(UW_avg)) %>%
  mutate(log.UW_avg.anom = scale(UW_avg)) -> UW36N_summer

UW36N_summer_Dc <- filter(UW36N_summer, 
                          (year >= min(Dc_nests$Year)-1) & 
                            (year <= max(Dc_nests$Year)-1))

climate.JM.lm.df <- data.frame(year = Dc_nests$Year,
                               counts = Dc_nests$Jamursba_Medi,
                               PDO_lag22 = PDO_Dc$avg,
                               UW36N_UWavg = UW36N_summer_Dc$UW_avg,
                               UW36N_log.UW_anom = UW36N_summer_Dc$log.UW_avg.anom)

climate.JM.lm.df <- mutate(climate.JM.lm.df,
                           log.counts = log(counts),
                           log.anomaly = scale(log.counts),
                           PDO_year = PDO_Dc$Year) %>%
  na.omit()

climate.JM.lm <- lm(log.anomaly ~ PDO_lag22 + UW36N_log.UW_anom, 
                    data = climate.JM.lm.df)
summary(climate.JM.lm)
```

I was expecting this sort of results... let's see what the plot shows.

```{r}
climate.JM.lm.df$fit <- predict(climate.JM.lm)
ggplot(data = climate.JM.lm.df) +
  geom_col(aes(x = year, y = log.anomaly)) +
  geom_line(aes(x = year, y = PDO_lag22),
            color = 'gray') +
  geom_line(aes(x = year, y = fit),
            color = 'blue')


```


