---
title: "Model comparison for data imputation"
output: html_notebook
---

```{r}
library(ggplot2)
library(bayesplot)
library(tidyverse)
library(lubridate)

```


This document and R code are used to describe how models were compared for imputation of missing data in nest counts of leatherback turtles at Jamusrba-Medi and Wermon. 

#Jamursba-Medi
For JM, the raw data looked like the following.

```{r}
# get JM data first:
data.0.JM <- read.csv('data/JM_nests.csv')

# create regularly spaced time series:
data.2.JM <- data.frame(Year = rep(min(data.0.JM$Year_begin,
                                       na.rm = T):max(data.0.JM$Year_begin,
                                                      na.rm = T),
                                   each = 12),
                        Month_begin = rep(1:12,
                                          max(data.0.JM$Year_begin,
                                              na.rm = T) -
                                            min(data.0.JM$Year_begin,
                                                na.rm = T) + 1)) %>%
  mutate(begin_date = as.Date(paste(Year,
                                    Month_begin,
                                    '01', sep = "-"),
                              format = "%Y-%m-%d"),
         Frac.Year = Year + (Month_begin-0.5)/12) %>%
  select(Year, Month_begin, begin_date, Frac.Year)

data.0.JM %>% mutate(begin_date = as.Date(paste(Year_begin,
                                                Month_begin,
                                                '01', sep = "-"),
                                          format = "%Y-%m-%d")) %>%
  mutate(Year = Year_begin,
         Month = Month_begin,
         f_month = as.factor(Month),
         f_year = as.factor(Year),
         Frac.Year = Year + (Month_begin-0.5)/12,
         Nests = JM.1) %>%
  select(Year, Month, Frac.Year, begin_date, Nests) %>%
  na.omit() %>%
  right_join(.,data.2.JM, by = "begin_date") %>%
  transmute(Year = Year.y,
            Month = Month_begin,
            Frac.Year = Frac.Year.y,
            Nests = Nests) %>%
  reshape::sort_df(.,vars = "Frac.Year") -> data.1.JM
# %>% filter(Year > 1998)

ggplot(data = data.1.JM) + geom_point(aes(x = Frac.Year, y = Nests)) + 
  geom_line(aes(x = Frac.Year, y = Nests))

```

Removing data prior to 1999...

```{r}
data.1.JM %>% filter(Year > 1998) -> data.1.JM
ggplot(data = data.1.JM) + geom_point(aes(x = Frac.Year, y = Nests)) + 
  geom_line(aes(x = Frac.Year, y = Nests))

```

The goal of the data imputation was to fill in the missing data. The basic model was autoregressive time-series model with time lag of 1 month. This decision was based on how counts were temporally correlated - there is a peak nesting month. 

```{r}
ggplot(data = data.1.JM) + geom_line(aes(x = Month, y = Nests, color = as.factor(Year)))
```


