---
title: "Test loo"
output: html_notebook
---

This script tests how the loo package performs on the data we use for the HILL analysis. Data are simulated using the fixed parameter values and the same models that were used in the analysis. Then, the same model is fitted to obtain posteriors on the parameters. This should provide the perfect pareto test values.

```{r}
rm(list=ls())
source('HILL_BiOp_functions.R')

library(jagsUI)
library(bayesplot)
library(dplyr)
library(loo)
```

Create data:
```{r}
Lnorm_Negbin <- readRDS('RData/SSAR1_pois_negbin_var_theta_JM_1999_jagsUI_2018-09-25.rds')

data.1 <- Lnorm_Negbin$data.1
means <- Lnorm_Negbin$jm$mean
theta.1 <- means$theta.1
theta.2 <- means$theta.2
sigma.1 <- means$sigma.pro1
sigma.2 <- means$sigma.pro2
r <- means$r
mu <- means$mu

m <- Lnorm_Negbin$jags.data$m

set.seed(123)
X <- p <- y <- vector(mode = "numeric", length = length(m))
X[1] <- mu
y[1] <- rnbinom(1, size = r, prob = r/(r + exp(X[1])))

for (k in 2:length(m)){
  theta <- ifelse(m[k] < 8, theta.1, theta.2)
  sigma <- ifelse(m[k] < 10 & m[k] > 4, sigma.1, sigma.2)
  X.mean <- theta * X[k-1]
  X[k] <- rlnorm(1, meanlog = X.mean, sdlog = sigma)
  
  lambda <- exp(X[k])
  p[k] <- r/(r + lambda)
  y[k] <- rnbinom(1, size = r, prob = p[k])
}

```

Then analyze the data and run "loo".

```{r}
if (!file.exists("RData/RData/SSAR1_lnorm_negbin_loo_test.rds")){
  MCMC.n.chains <- 5
  MCMC.n.samples <- 250000
  MCMC.n.burnin <- 150000
  MCMC.n.thin <- 50
  
  MCMC.params <- list(n.chains = MCMC.n.chains,
                      n.samples = MCMC.n.samples,
                      n.burnin = MCMC.n.burnin,
                      n.thin = MCMC.n.thin)
  
  jags.data <- list(y = y,
                    m = m,
                    T = length(m))
  
  #load.module('dic')
  jags.params <- c('theta.1', "theta.2", 
                   'sigma.pro1', 'sigma.pro2', "r",
                   'mu', 'X', 'deviance', 'loglik')
  
  jm <- jags(jags.data,
             inits = NULL,
             parameters.to.save= jags.params,
             model.file = 'models/model_SSAR1_lnorm_negbin_var_theta.txt',
             n.chains = MCMC.n.chains,
             n.burnin = MCMC.n.burnin,
             n.thin = MCMC.n.thin,
             n.iter = MCMC.n.samples,
             DIC = T, parallel=T)
  
  #g.diag1 <- gelman.diag(jm$samples)
  Rhat <- jm$Rhat
  
  loo.out <- pareto.k.diag(jm, MCMC.params, jags.data)
  
  # extract Xs - the state model
  Xs.stats <- data.frame(low_X = jm$q2.5$X,
                         median_X = jm$q50$X,
                         high_X = jm$q97.5$X,
                         time = data.1$Frac.Year,
                         obsY = y,
                         month = m,
                         year = data.1$Year)
  
  Xs.year <- group_by(Xs.stats, year) %>% summarize(median = sum(median_X),
                                                    low = sum(low_X),
                                                    high = sum(high_X))
  
  
  results<- list(jags.data = jags.data,
                 Xs.stats = Xs.stats,
                 MCMC.params = MCMC.params,
                 Rhat = Rhat,
                 jm = jm,
                 loo.out = loo.out)
  
  saveRDS(results,
          file = paste0('RData/SSAR1_lnorm_negbin_loo_test.rds'))
} else {
  results <- readRDS(file = "RData/SSAR1_lnorm_negbin_loo_test.rds")
}

```

and look at the results

```{r}
Xs.stats <- results$Xs.stats

p.1 <- ggplot() +
  #geom_point(data = ys.stats,
  #           aes(x = time, y = mode_y), color = "blue") +
  #geom_line(data = Xs.stats,
  #          aes(x = time, y = mode_X), color = 'blue') +
  geom_line(data = Xs.stats,
            aes(x = time, y = exp(high_X)), color = "red",
            linetype = 2) +
  geom_point(data = Xs.stats,
             aes(x = time, y = exp(median_X)), color = "red",
             alpha = 0.5) +
  geom_line(data = Xs.stats,
            aes(x = time, y = exp(median_X)), color = "red",
            alpha = 0.5) +
  geom_point(data = Xs.stats,
             aes(x = time, y = y), color = "green",
             alpha = 0.5)

base_theme <- ggplot2::theme_get()
library(bayesplot)

# set back to the base theme:
ggplot2::theme_set(base_theme)
mcmc_trace(jm$samples, c('theta.1', "theta.2", "mu",
                         "sigma.pro1", "sigma.pro2", "r"))
mcmc_dens(jm$samples, c('theta.1', "theta.2", "mu",
                        "sigma.pro1", "sigma.pro2", "r"))

```

