---
title: "Negative binomial distribution in R and jags"
output: html_notebook
---

This document compares how negative binomial distribution is defined in R and jags. Specifically, when the distribution is defined in terms of mean (mu) and "dispersion" parameter (r; it is also called "size" parameter), the probability of occurrence (p) is defined as the probability of "success" in R and jags, even though the density is computed for the "failures." This is the standard in defining negative binomial distribution but seems confusing... 

In the R help document, it is stated that "An alternative parametrization (often used in ecology) is by the mean mu (see above), and size, the dispersion parameter, where prob = size/(size+mu). The variance is mu + mu^2/size in this parametrization."  The mean (mu) is defined as n(1-p)/p, where the density of the distribution with size = n and probability = p is defined as p(x|n, p) = (G(x+n)/(G(n) x!))(p^n)((1-p)^x), x = 0, 1, ..., n and 0 < p <= 1.  G denotes the gamma function. I'm not sure if the upper bound of x should be n.  In the independent Bernoulli trial sense, this equation states that x failures and n successes, where p is defined as the probability of "success." However, the density is computed for the x failures. So, for example, probability of # failures = 5, given # success = 2 and prob of success = 0.4 is

```{r}
dnbinom(x = 5, size = 2, prob = 0.4)
```

The same density should be computes using the definition

```{r}
x <- 5
n <- 2
p <- 0.4
(gamma(x + n)/(gamma(n) * factorial(x))) * (p^n) * ((1 - p)^x)
```

In this example, n < x. So, the upper bound of x does not have to be n. (error in R help) The important point here is that the probability (p) is defined with respect to "success", which goes with the "size" parameter (a.k.a. dispersion parameter). 

In jags, this is also the case.

```{r}
library(jagsUI)
MCMC.n.chains <- 1
MCMC.n.samples <- 10
MCMC.n.burnin <- 5
MCMC.n.thin <- 1

jags.data <- list(x = 5,
                  p = 0.4,
                  r = 2)

jags.params <- c('loglik')

jags.model <- cat("model{
    loglik <- logdensity.negbin(x, p, r)

}", file = "jags_model_negbin.txt")

jm <- jags(jags.data,
           inits = NULL,
           parameters.to.save= jags.params,
           model.file = 'jags_model_negbin.txt',
           n.chains = MCMC.n.chains,
           n.burnin = MCMC.n.burnin,
           n.thin = MCMC.n.thin,
           n.iter = MCMC.n.samples,
           DIC = T, parallel=T)
```

We can compute the density from the log-likelihood.

```{r}
exp(jm$samples[[1]][1,])
```

This is identical to the result from R... 

So... what happens here is the probability (p) is defined as the probability of "success".

In Wikipedia, the probability is defined as the probability of "failure", which makes more sense... 

