---
title: "Nest count estimations and trend for Western Pacific leatherback nesting beaches"
output: html_notebook
---

This analysis tries to fill in missing nest counts data. A parametric model is fitted to years with complete data. The estimated parameters are used to fill in years without complete data. Ultimately, it aims to look at how the counts are changing over time. 

This analysis estimates trends of leatherback turtle nests at two Indonesia nesting beaches (Jamursba-Medi and Wermon) using the annual nest counts. Raw data for the analysis are the recorded number of nests per month. Missing data are imputed using auto-regressive time-series analysis. A few models are used to compare performance of imputation and one with the narrowest prediction (imputation) intervals is used for the final analysis. 

```{r echo=FALSE} 
# load some libraries and functions. 
rm(list=ls())
source('HILL_BiOp_functions.R')
library(lme4)
library(mgcv)
library(rjags)
#library(loo)
base_theme <- ggplot2::theme_get()
library(bayesplot)
#
# # set back to the base theme:
ggplot2::theme_set(base_theme)
```

#Jamursba-Medi
First, data for Jamursba-Medi are analyzed.  
```{r}
# get JM data first:
data.0.JM <- read.csv('data/JM_nests.csv')

# create regularly spaced time series:
data.2.JM <- data.frame(Year = rep(min(data.0.JM$Year_begin, 
                                       na.rm = T):max(data.0.JM$Year_begin, 
                                                      na.rm = T),
                                   each = 12),
                        Month_begin = rep(1:12, 
                                          max(data.0.JM$Year_begin, 
                                              na.rm = T) - 
                                            min(data.0.JM$Year_begin, 
                                                na.rm = T) + 1)) %>%
  mutate(begin_date = as.Date(paste(Year,
                                    Month_begin,
                                    '01', sep = "-"),
                              format = "%Y-%m-%d"),
         Frac.Year = Year + (Month_begin-0.5)/12) %>%
  select(Year, begin_date, Frac.Year)

# create time-duration filed (in yrs)
# define dates with begin and end dates:
data.0.JM %>% mutate(begin_date = as.Date(paste(Year_begin,
                                                Month_begin,
                                                '01', sep = "-"),
                                          format = "%Y-%m-%d")) %>%
  mutate(end_day = if_else(Month_end.1 == 1 | Month_end.1 == 3 | Month_end.1 == 5 |
                             Month_end.1 == 7 | Month_end.1 == 8 | Month_end.1 == 10 |
                             Month_end.1 == 12, 31,
                           if_else(Month_end.1 == 2,
                                   if_else(leap_year(Year_end) == TRUE,
                                           29, 28), 30))) %>%
  mutate(end_date = as.Date(paste(Year_end, Month_end.1, end_day,
                                  sep = "-"),
                            format = "%Y-%m-%d")) %>%
  mutate(cumu_days = as.numeric(end_date - begin_date[1] + 1),
         num_days = as.numeric(end_date - begin_date) + 1,
         Year = Year_begin,
         Month = Month_begin,
         f_month = as.factor(Month),
         f_year = as.factor(Year),
         Nests = JM.1)   %>%
  select(Year, Month, begin_date,
         end_date, num_days, cumu_days,
         f_month, f_year, 
         Nests)  %>%  
  na.omit() %>%
  right_join(.,data.2.JM, by = "begin_date") %>%
  reshape::sort_df(.,vars = "Frac.Year") -> data.1.JM

```

Raw data for Jamursba-Medi indicate that data were pooled over a few months prior to 2011. These pooled data can be used later when looking at the long-term trend but not useful for filling in the more recent missing data. 

```{r}
p.JM.all <- ggplot(data.1.JM) + 
  geom_point(aes(x = Frac.Year, y = Nests)) + 
  geom_line(aes(x = Frac.Year, y = Nests)) +
  scale_x_continuous(breaks = seq(1980, 2020, 5),
                     limits = c(1980, 2020)) +
  labs(x = '', y = '# nests', title = "Jamursba-Medi")  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))

p.JM.all
if (save.fig) ggsave(filename = 'figures/JM_all.png',
                     plot = p.JM.all,
                     dpi = 600, height = 6, width = 8, units = "in")
```


No data were collected in 1998. Since then at least some data were collected within each year. So, we look at the data since 1999:

```{r}
data.1.JM.1999 <- filter(data.1.JM, Year.y > 1998)
p.JM.1999 <- ggplot(data.1.JM.1999) + 
  geom_point(aes(x = Frac.Year, y = Nests)) + 
  geom_line(aes(x = Frac.Year, y = Nests)) +
  scale_x_continuous(breaks = seq(1999, 2019, 5),
                     limits = c(1999, 2019)) +
  labs(x = '', y = '# nests', 
       title = "Jamursba-Medi")  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))

p.JM.1999
if (save.fig) ggsave(filename = 'figures/JM_1999.png',
                     plot = p.JM.1999,
                     dpi = 600, height = 6, 
                     width = 8, units = "in")
```

First, I look at the time series in the natural log scle.

```{r}
data.1.JM.1999$log_Nests <- log(data.1.JM.1999$Nests)
p.log.JM.1999 <- ggplot(data.1.JM.1999) + 
  geom_point(aes(x = Frac.Year, y = log_Nests)) + 
  geom_line(aes(x = Frac.Year, y = log_Nests)) +
  scale_x_continuous(breaks = seq(1999, 2019, 5),
                     limits = c(1999, 2019)) +
  labs(x = '', y = 'log(# nests)', 
       title = "Jamursba-Medi")  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))

p.log.JM.1999
if (save.fig) ggsave(filename = 'figures/JM_log_1999.png',
                     plot = p.log.JM.1999,
                     dpi = 600, height = 6, 
                     width = 8, units = "in")
```

To fill in (or estiamte) the missing data points, we fit a simple auto-regressive model to the data. The analysis was done in a different script and results saved in a .RData file. 

```{r}
load("RData/SSAR1_JM_logN_1999_2018-06-07.RData")
results.JM_SSAR1_logN$jm
```

Data used for this analysis:

```{r}
results.JM_SSAR1_logN$bugs.data
```

The MCMC run converged fine according to the Gelman diagnostic.

```{r}
results.JM_SSAR1_logN$g.diag
```

Visual inspection of convergence of the process parameters. They all look good, except sigma.obs.
```{r}
zm.JM.logN.1999 <- results.JM_SSAR1_logN$zm
mcmc_trace(zm.JM.logN.1999, 
           c("sigma.pro", "sigma.obs", 
             "theta", "mu"))
```

Shows in density plots too. 
```{r}
mcmc_dens(zm.JM.logN.1999, 
           c("sigma.pro", "sigma.obs", 
             "theta", "mu"))

```

I look at the predicted values as a time-series plot. Red is the estimated missing values and green are observed counts. 5 and 95% confidence limits are shown in dashed line. Although I called mode_X, they should have been median_X because they were 50 percentiles... oops. 

```{r warning=F}
Xs.stats <- results.JM_SSAR1_logN$Xs.stats
ys.stats <- results.JM_SSAR1_logN$ys.stats
ys.stats$log_obsY <- log(results.JM_SSAR1_logN$data.1$Nests)

p.JM.logN.predicted.1999 <- ggplot() +
  #geom_point(data = ys.stats,
  #           aes(x = time, y = mode_y), color = "blue") +
  #geom_line(data = Xs.stats,
  #          aes(x = time, y = mode_X), color = 'blue') +
  geom_line(data = Xs.stats,
            aes(x = time, y = exp(high_X)), color = "red",
            linetype = 2) +
  geom_point(data = Xs.stats,
             aes(x = time, y = exp(mode_X)), color = "red",
             alpha = 0.5, size = 2) +
  geom_line(data = Xs.stats,
            aes(x = time, y = exp(mode_X)), color = "red",
            alpha = 0.5) +
  geom_line(data = Xs.stats,
            aes(x = time, y = exp(low_X)), color = "red",
            linetype = 2) +
  geom_point(data = ys.stats,
             aes(x = time, y = log_obsY), 
             color = "green",
             alpha = 0.5) +
  scale_y_continuous(limits = c(0, 12)) +
  labs(x = '', y = "log(Nest counts)",
       title = "Jamursba-Medi")

p.JM.logN.predicted.1999
if (save.fig) ggsave(filename = 'figures/JM_logN_predicted_1999.png', 
                     plot = p.JM.logN.predicted.1999,
                     dpi = 600, height = 6, 
                     width = 8, units = "in")
```

The missing values are basically linearly interpolated between observed values. Converting them back into the non-log scale:

```{r}
p.JM.N.predicted.1999 <- ggplot() +
  #geom_point(data = ys.stats,
  #           aes(x = time, y = mode_y), color = "blue") +
  #geom_line(data = Xs.stats,
  #          aes(x = time, y = mode_X), color = 'blue') +
  geom_line(data = Xs.stats,
            aes(x = time, y = exp(exp(high_X))), color = "red",
            linetype = 2) +
  geom_point(data = Xs.stats,
             aes(x = time, y = exp(exp(mode_X))), color = "red",
             alpha = 0.5, size = 2) +
  geom_line(data = Xs.stats,
            aes(x = time, y = exp(exp(mode_X))), color = "red",
            alpha = 0.5) +
  geom_line(data = Xs.stats,
            aes(x = time, y = exp(exp(low_X))), color = "red",
            linetype = 2) +
  geom_point(data = ys.stats,
             aes(x = time, y = obsY), 
             color = "green",
             alpha = 0.5) +
  scale_y_continuous(limits = c(0, 10000)) +
  labs(x = '', y = "Nest counts",
       title = "Jamursba-Medi")

p.JM.N.predicted.1999
```

The upper bound is not useful... So, looking at these in the log space may not be helpful. I think the change in variance of nest counts between summer and winter is an important biological signal and should be modeled. The variability is obvious when the observed nest counts are plotted monthly:

```{r warning=FALSE}
p.JM.monthly.1999 <- ggplot(data.1.JM.1999) + 
  geom_point(aes(x = Month, y = Nests, color = f_year)) + 
  geom_line(aes(x = Month, y = Nests, color = f_year)) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  labs(x = 'Month', y = '# nests', 
       title = "Jamursba-Medi", color = "Year")  +
  theme(axis.text = element_text(size = 12),
        text = element_text(size = 12))
#        legend.position = c(0.9, 0.6))

p.JM.monthly.1999
if (save.fig) ggsave(filename = 'figures/JM_monthly_1999.png', 
                     plot = p.JM.monthly.1999,
                     dpi = 600, height = 6, 
                     width = 8, units = "in")
```

Using the cyclic nature of nesting, we should be able to fit a model to fill in the missing data points. We need to make some assumptions about how data were collected, unless we have a measure of effort every year.  I ran time-series models to compare how they do in terms of filling in the missing data points. 

First, I try to fill in the missing data points using an AR(1) state-space model and data since 1999. JAGS model for this analysis:

```{r}
load("RData/SSAR1_month_var_JM_1999_2018-06-07.RData")
results.JM_SSAR1_month_var$jm
```

It has two process standard deviations; one for summer (May through September; sigma.pro1) and another for winter (sigma.pro2). Observation has only one standard deviation under the assumption that it does not change with time. Data for this analysis is the same as before so not shown here. Convergence seemed to reach, according to the Gelman diagnostic, except the standard deviation of the observation.

```{r}
results.JM_SSAR1_month_var$g.diag
```

Posterior trace plots also show the same results.  

```{r}
zm.var.JM.1999 <- results.JM_SSAR1_month_var$zm
mcmc_trace(zm.var.JM.1999, 
           c("sigma.pro1", "sigma.pro2", 
             "theta", "sigma.obs"))
```

  
Posterior density plots of the same parameters:

```{r}
zm.var.JM.1999 <- results.JM_SSAR1_month_var$zm
mcmc_dens(zm.var.JM.1999, 
          c("sigma.pro1", "sigma.pro2", 
            "theta", "sigma.obs"))
```


Predicted missing values from this model:
```{r warning=F}
Xs.stats <- results.JM_SSAR1_month_var$Xs.stats
ys.stats <- results.JM_SSAR1_month_var$ys.stats

p.JM.month_var.predicted.1999 <- ggplot() +
  #geom_point(data = ys.stats,
  #           aes(x = time, y = mode_y), color = "blue") +
  #geom_line(data = Xs.stats,
  #          aes(x = time, y = mode_X), color = 'blue') +
  geom_line(data = Xs.stats,
            aes(x = time, y = high_X), color = "red",
            linetype = 2) +
  geom_point(data = Xs.stats,
             aes(x = time, y = median_X), color = "red",
             alpha = 0.5, size = 2) +
  geom_line(data = Xs.stats,
            aes(x = time, y = median_X), color = "red",
            alpha = 0.5) +
  geom_line(data = Xs.stats,
            aes(x = time, y = low_X), color = "red",
            linetype = 2) +
  geom_point(data = ys.stats,
             aes(x = time, y = obsY), 
             color = "green",
             alpha = 0.5) +
  #scale_y_continuous(limits = c(0, 5000)) +
  labs(x = '', y = "Nest counts",
       title = "Jamursba-Medi")

p.JM.month_var.predicted.1999
if (save.fig) ggsave(filename = 'figures/JM_month_var_predicted_1999.png', 
                     plot = p.JM.logN.predicted.1999,
                     dpi = 600, height = 6, 
                     width = 8, units = "in")
```

It looks a lot better than the previous model but can we make it even better?

The next model has two variances and two "slope" parameters. As before, two variances correspond to nesting (May, June, July, August, and September; sigma.pro1) and non-nesting (sigma.pro2) seasons, whereas two slopes are for increasing (Jan to Jul) and decreasing (Aug to Dec) phases of each year.  Because it takes a bit of time to run these, output has been saved as an RData file. 

First, JAGS model for the analysis:
```{r}
load("RData/SSAR1_month_JM_var_theta_1999_2018-06-06.RData")
results.JM_SSAR1_month_var_theta$jm
```

Data used for this analysis:

```{r}
results.JM_SSAR1_month_var_theta$bugs.data
```

The MCMC run converged fine according to the Gelman diagnostic.

```{r}
results.JM_SSAR1_month_var_theta$g.diag
```

Visual inspection of convergence of the process parameters. They all look good.
```{r}
zm.var.theta.JM.1999 <- results.JM_SSAR1_month_var_theta$zm
mcmc_trace(zm.var.theta.JM.1999, 
           c("sigma.pro1", "sigma.pro2", 
             "theta.1", "theta.2"))
```

The observation variance parameter did not converge well. 

```{r}
mcmc_trace(zm.var.theta.JM.1999, "sigma.obs")
```

Shows in density plot too. 
```{r}
mcmc_dens(zm.var.theta.JM.1999, 
          "sigma.obs")

```

I look at the predicted values as a time-series plot. Red is the estimated missing values and green are observed counts. 5 and 95% confidence limits are shown in dashed line. Although I called mode_X, they should have been median_X because they were 50 percentiles... oops. 

```{r warning=F}
Xs.stats <- results.JM_SSAR1_month_var_theta$Xs.stats
ys.stats <- results.JM_SSAR1_month_var_theta$ys.stats
ys.stats$obsY <- results.JM_SSAR1_month_var_theta$data.1$Nests

p.JM.predicted_month_var_theta_1999 <- ggplot() +
  #geom_point(data = ys.stats,
  #           aes(x = time, y = mode_y), color = "blue") +
  #geom_line(data = Xs.stats,
  #          aes(x = time, y = mode_X), color = 'blue') +
  geom_line(data = Xs.stats,
            aes(x = time, y = high_X), color = "red",
            linetype = 2) +
  geom_point(data = Xs.stats,
             aes(x = time, y = mode_X), color = "red",
             alpha = 0.5, size = 2) +
  geom_line(data = Xs.stats,
            aes(x = time, y = mode_X), color = "red",
            alpha = 0.5) +
  geom_line(data = Xs.stats,
            aes(x = time, y = low_X), color = "red",
            linetype = 2) +
  geom_point(data = ys.stats,
             aes(x = time, y = obsY), color = "green",
             alpha = 0.5) +
  labs(x = '', y = "Nest counts",
       title = "Jamursba-Medi")

p.JM.predicted_month_var_theta_1999
if (save.fig) ggsave(filename = 'figures/JM_predicted_month_var_theta_1999.png', 
                     plot = p.JM.predicted.1999,
                     dpi = 600, height = 6, 
                     width = 8, units = "in")
```

The uncertainty shrunk quite a bit with this model. We may go with this one? 


Sum within each year to find change in trend over the years.
```{r}
#sum.posterior is in Dc_Indonesia_nesting_fcn.R

yrs <- as.matrix(c(min(Xs.stats$year):max(Xs.stats$year)))
X.posteriors.all <- lapply(apply(yrs, 
                                 MARGIN = 1,
                                 FUN = sum.posterior, 
                                 Xs.stats = Xs.stats, 
                                 zm = zm.var.theta.JM.1999),
                           FUN = function(x) rowSums(x$samples))

Xs.year <- as.data.frame(matrix(unlist(lapply(X.posteriors.all, 
                                              FUN = quantile,
                                              c(0.025, 
                                                0.5, 
                                                0.975))),
                                ncol = 3, byrow = T)) 
colnames(Xs.year) <- c("low", "total", "high")
Xs.year <-mutate(Xs.year, 
                 year = c(min(Xs.stats$year):max(Xs.stats$year)))

# Xs.stats %>% #mutate(f.year = as.factor(year)) %>%
#   group_by(year) %>% summarise(total = sum(mode_X), 
#                                low = sum(low_X),
#                                high = sum(high_X)) -> Xs.year

p.JM.estimated.counts <- ggplot(Xs.year) + 
  geom_point(aes(x = year, y = total))+
  geom_errorbar(aes(x = year, ymin = low, ymax = high)) + 
  labs(title = 'Jamursba-Medi', x = '', y = "Nest counts")

if (save.fig) ggsave(filename = 'figures/JM_estimated_counts.png', 
                     plot = p.JM.estimated.counts,
                     dpi = 600, height = 6, 
                     width = 8, units = "in")
p.JM.estimated.counts
```

Adding the earlier year's data as just sum of counts:

```{r}
data.1.JM %>% group_by(Year.y) %>%
  summarise(sum_Nests = sum(Nests, na.rm = T),
            year = mean(Year.y, na.rm = T),
            effort = sum(num_days, na.rm = T))%>%
  filter(effort > 0) -> data.1.JM.ByYear

p.JM.estimated.observed <- p.JM.estimated.counts + 
  geom_point(data = data.1.JM.ByYear,
             aes(x = year, y = sum_Nests, size = effort),
             color = "red", alpha = 0.5)
p.JM.estimated.observed
```

The size of circles indicates effort in the number of sampling days. 




#Winter vs. Summer nest trends
Because of the  nearshore currents are different between summer and winter, it may make more sense to look at trends for winter and summer counts separately. So, we have to pull in the predicted numbers from model outputs. First from Wermon. 

```{r}
# get Wermon output
load("RData/SSAR1_month_W_var_theta_2006to2017_2018-05-07.RData")
zm.W <- results.W_SSAR1_month_var_theta$zm
Xs.stats.W <- results.W_SSAR1_month_var_theta$Xs.stats
ys.stats.W <- results.W_SSAR1_month_var_theta$ys.stats

# for this analysis, we need to extract just summer/winter numbers.
yrs.W <- as.matrix(c(min(Xs.stats.W$year):max(Xs.stats.W$year)))
# summer is May through September
X.posteriors.all.W.summer <- lapply(apply(yrs.W, 
                                          MARGIN = 1,
                                          FUN = sum.posterior, 
                                          months = c(5:9),
                                          Xs.stats = Xs.stats.W, 
                                          zm = zm.W),
                                    FUN = function(x) rowSums(x$samples))

Xs.year.W.summer <- as.data.frame(matrix(unlist(lapply(X.posteriors.all.W.summer, 
                                                       FUN = quantile,
                                                       c(0.025, 0.5, 0.975))),
                                         ncol = 3, byrow = T)) 

colnames(Xs.year.W.summer) <- c("low", "total", "high")
Xs.year.W.summer <- mutate(Xs.year.W.summer, 
                           year = c(min(Xs.stats.W$year):max(Xs.stats.W$year)))

# winter is Jan-Apr and October-Dece
X.posteriors.all.W.winter <- lapply(apply(yrs.W, 
                                          MARGIN = 1,
                                          FUN = sum.posterior, 
                                          months = c(1:4, 10:12),
                                          Xs.stats = Xs.stats.W, 
                                          zm = zm.W),
                                    FUN = function(x) rowSums(x$samples))

Xs.year.W.winter <- as.data.frame(matrix(unlist(lapply(X.posteriors.all.W.winter, 
                                                       FUN = quantile,
                                                       c(0.025, 0.5, 0.975))),
                                         ncol = 3, byrow = T)) 
colnames(Xs.year.W.winter) <- c("low", "total", "high")
Xs.year.W.winter <- mutate(Xs.year.W.winter, 
                           year = c(min(Xs.stats.W$year):max(Xs.stats.W$year)))
```

Next is Jamursba-Medi:
```{r}
load("RData/SSAR1_month_JM_var_theta_1999_2018-05-03.RData")
zm.JM <- results.JM_SSAR1_month_var_theta$zm
Xs.stats.JM <- results.JM_SSAR1_month_var_theta$Xs.stats
ys.stats.JM <- results.JM_SSAR1_month_var_theta$ys.stats

# for this analysis, we need to extract just summer/winter numbers.
yrs.JM <- as.matrix(c(min(Xs.stats.JM$year):max(Xs.stats.JM$year)))
# summer is May through September
X.posteriors.all.JM.summer <- lapply(apply(yrs.JM, 
                                           MARGIN = 1,
                                           FUN = sum.posterior, 
                                           months = c(5:9),
                                           Xs.stats = Xs.stats.JM, 
                                           zm = zm.JM),
                                     FUN = function(x) rowSums(x$samples))

Xs.year.JM.summer <- as.data.frame(matrix(unlist(lapply(X.posteriors.all.JM.summer, 
                                                        FUN = quantile,
                                                        c(0.025, 0.5, 0.975))),
                                          ncol = 3, byrow = T)) 
colnames(Xs.year.JM.summer) <- c("low", "total", "high")
Xs.year.JM.summer <- mutate(Xs.year.JM.summer, 
                            year = c(min(Xs.stats.JM$year):max(Xs.stats.JM$year)))

# winter is Jan-Apr and October-Dece
X.posteriors.all.JM.winter <- lapply(apply(yrs.JM, 
                                           MARGIN = 1,
                                           FUN = sum.posterior, 
                                           months = c(1:4, 10:12),
                                           Xs.stats = Xs.stats.JM, 
                                           zm = zm.JM),
                                     FUN = function(x) rowSums(x$samples))

Xs.year.JM.winter <- as.data.frame(matrix(unlist(lapply(X.posteriors.all.JM.winter, 
                                                        FUN = quantile,
                                                        c(0.025, 0.5, 0.975))),
                                          ncol = 3, byrow = T)) 
colnames(Xs.year.JM.winter) <- c("low", "total", "high")
Xs.year.JM.winter <- mutate(Xs.year.JM.winter, 
                            year = c(min(Xs.stats.JM$year):max(Xs.stats.JM$year)))
```

Then combine them together. If there was no data in either location (NA), I put zero. 

```{r}
Xs.year.JM.summer %>% full_join(Xs.year.W.summer, by = "year") %>%
  transmute(year = year, 
            low = low.x + ifelse(is.na(low.y), 0, low.y), 
            total = total.x + ifelse(is.na(total.y), 0, total.y), 
            high = high.x + ifelse(is.na(high.y), 0, high.y)) -> Xs.year.summer

p.summer.log <- ggplot(data = Xs.year.summer) + 
  geom_point(aes(x = year, y = log(total)))+
  geom_errorbar(aes(x = year, ymin = log(low), ymax = log(high))) + 
  labs(title = 'Summer (May-Sep)', x = '', y = "log(Nest counts)")

if (save.fig) ggsave(filename = 'figures/summer_JMandW_log.png', 
                     plot = p.summer.log,
                     dpi = 600, height = 6, 
                     width = 8, units = "in")
p.summer.log
```

In non-log scale:

```{r}
p.summer <- ggplot(data = Xs.year.summer) + 
  geom_point(aes(x = year, y = total))+
  geom_errorbar(aes(x = year, ymin = low, ymax = high)) + 
  labs(title = 'Summer (May-Sep)', x = '', y = "Nest counts")

if (save.fig) ggsave(filename = 'figures/summer_JMandW.png', 
                     plot = p.summer.log,
                     dpi = 600, height = 6, 
                     width = 8, units = "in")
p.summer
```

Even in the non-log scale, the increase in the last five years seems obvious... How do we capture this for predicting the future trend? 

Fit a GAM:

```{r}
Xs.year.summer %>% mutate(year.0 = year - min(year)) -> Xs.year.summer
fit.gam.summer <- gam(total ~ s(year.0), 
                      data = Xs.year.summer)
sum.gam.summer <- summary(fit.gam.summer)
sum.gam.summer
```

Then make the response curve.

```{r}
gam.val.summer <- plot(fit.gam.summer)
gam.val.summer.df <- data.frame(x = gam.val.summer[[1]]$x,
                                fit = gam.val.summer[[1]]$fit,
                                se = gam.val.summer[[1]]$se)

Xs.year.summer %>% 
  mutate(log.total = log(total)) %>%
  mutate(log.total.std = log(total) - mean(log(total), na.rm = T)) %>%
  mutate(total.std = total - mean(total, na.rm = T)) -> Xs.year.summer3

p.summer.GAM <- ggplot() + 
  geom_line(data = gam.val.summer.df,
            aes(x = x+min(Xs.year.summer$year), 
                y = fit)) + 
  geom_line(data = gam.val.summer.df,
            aes(x = x+min(Xs.year.summer$year), 
                y = fit + 2*se), linetype = 2) + 
  geom_line(data = gam.val.summer.df,
            aes(x = x+min(Xs.year.summer$year), 
                y = fit - 2*se), linetype = 2) + 
  geom_point(data = Xs.year.summer3,
             aes(x = year, y = total.std),
             color = 'red') + 
  labs(title = 'Summer (May-Sep)', x = '', 
       y = paste0('GAM(', signif(sum.gam.summer$edf, digits = 3), ')'))
  

if (save.fig) ggsave(filename = 'figures/summer_GAM.png', 
                     plot = p.summer.GAM,
                     dpi = 600, height = 6, 
                     width = 8, units = "in")
p.summer.GAM
```

Note that trying to extrapolate the "trend" into the future is not useful because the prediction is depending on what we have observed in the past. What have others done? The green turtle SRT approach was not useful because it just created trajectories based on estimated growth rate and its uncertainty.  

```{r}
Xs.year.JM.winter %>% full_join(Xs.year.W.winter, by = "year") %>%
  transmute(year = year, 
            low = low.x + ifelse(is.na(low.y), 0, low.y), 
            total = total.x + ifelse(is.na(total.y), 0, total.y), 
            high = high.x + ifelse(is.na(high.y), 0, high.y)) -> Xs.year.winter

p.winter.log <- ggplot(data = Xs.year.winter) + 
  geom_point(aes(x = year, y = log(total)))+
  geom_errorbar(aes(x = year, 
                    ymin = log(low), 
                    ymax = log(high))) + 
  labs(title = 'Winter (Oct-Apr)', x = '', 
       y = "log(Nest counts)")

if (save.fig) ggsave(filename = 'figures/winter_JMandW_log.png', 
                     plot = p.winter.log,
                     dpi = 600, height = 6, 
                     width = 8, units = "in")
p.winter.log
```

In non-log scale:

```{r}
p.winter <- ggplot(data = Xs.year.winter) + 
  geom_point(aes(x = year, y = total))+
  geom_errorbar(aes(x = year, 
                    ymin = low, 
                    ymax = high)) + 
  labs(title = 'Winter (Oct-Apr)', x = '', 
       y = "Nest counts")

if (save.fig) ggsave(filename = 'figures/winter_JMandW.png', 
                     plot = p.summer.log,
                     dpi = 600, height = 6, 
                     width = 8, units = "in")
p.winter
```

Fit a GAM:

```{r}
Xs.year.winter %>% mutate(year.0 = year - min(year)) -> Xs.year.winter
fit.gam.winter <- gam(total ~ s(year.0), 
                      data = Xs.year.winter)
sum.gam.winter <- summary(fit.gam.winter)
sum.gam.winter
```

Then make the response curve.

```{r}
gam.val.winter <- plot(fit.gam.winter)
gam.val.winter.df <- data.frame(x = gam.val.winter[[1]]$x,
                                fit = gam.val.winter[[1]]$fit,
                                se = gam.val.winter[[1]]$se)

Xs.year.winter %>% 
  mutate(log.total = log(total)) %>%
  mutate(log.total.std = log(total) - mean(log(total), na.rm = T)) %>%
  mutate(total.std = total - mean(total, na.rm = T)) -> Xs.year.winter3

p.winter.GAM <- ggplot() + 
  geom_line(data = gam.val.winter.df,
            aes(x = x+min(Xs.year.summer$year), 
                y = fit)) + 
  geom_line(data = gam.val.winter.df,
            aes(x = x+min(Xs.year.summer$year), 
                y = fit + 2*se), linetype = 2) + 
  geom_line(data = gam.val.winter.df,
            aes(x = x+min(Xs.year.summer$year), 
                y = fit - 2*se), linetype = 2) + 
  geom_point(data = Xs.year.winter3,
             aes(x = year, y = total.std),
             color = 'red') + 
  labs(title = 'Winter (Oct-Apr)', x = '', 
       y = paste0('GAM(', signif(sum.gam.winter$edf, digits = 3), ')'))
  

if (save.fig) ggsave(filename = 'figures/winter_GAM.png', 
                     plot = p.winter.GAM,
                     dpi = 600, height = 6, 
                     width = 8, units = "in")
p.winter.GAM
```